Review of Edge Detectors

### 2015-04-01 edge-detection-review-3rd-revision-v3.tex
finished correcting text

\newcommand{\demourl}{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}

@@

In the 70's, edge detection methods were implemented using small operators 
(such as Sobel masks), attempting to compute an approximation of the
first derivative of the image \cite{Gonzalez2007Digital}. Section \ref{sec:first} describes such algorithms 
and serves as an introduction to a more sophisticated 
analysis of the edge detection process.

% en este parrafo, puede ser cambiar trough por valley
In 1980 Marr and Hildreth \cite{segm:edge_region:marr:84:digital_step} argued that intensity changes are not independent 
of image scale, so edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or valley) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
Haralick's algorithm \cite{bb20239} is an alternative approach based on the second derivative 
which is also reviewed in Section \ref{sec:second}. This algorithm has the particularity of 
proposing a model to approximate the image locally around a point; then, using this model, 
an approximation to the second derivative of the image can be calculated analytically and 
finding edges is achieved by imposing a condition over the model parameters.

@@

\nocite{IPOL}

@@

The edge detection methods discussed in the previous section are simply based on filtering the 
image with different masks, without taking into account the characteristics of the edges or 
the noise in the image.
The algorithms that will presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:84:digital_step} and Haralick \cite{bb20239}) 
are based on the second derivative of the image, and take steps to reduce noise before 
detecting edges in the image.

@@

\begin{algorithm}[t!]
\caption{Marr-Hildreth edge detection algorithm.}
\label{algo:mh}
\begin{algorithmic}[1]
\Require $im$ input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
\If{Gaussian kernel}
	\State $kernel \leftarrow$ generate\_gaussian\_kernel($n$,$\sigma$)
	\State $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
	\State Define Laplacian operator $laplacian$ \Comment{See mask definition in Eq. \ref{eq:laplacian:mask}}
	\State $im_{LAPL} \leftarrow$ convolution($im_{SMOOTHED}$,$laplacian$)
\Else
\State $log\_kernel \leftarrow$ generate\_log\_kernel($n$,$\sigma$)
\State $im_{LAPL} \leftarrow$ convolution($im$,$log\_kernel$)
\EndIf
\State $max_L \leftarrow 0$
\ForAll{pixel $i$ in image $im_{LAPL}$}
	\If{$im_{LAPL}[i]>max_L$}
		\State $max_L \leftarrow im_{LAPL}[i]$
	\EndIf
\EndFor
\ForAll{pixel $i$ in image $im_{LAPL}$, except borders}
	\ForAll{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
		\If{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \bf{and} ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
			\State $edges[i] \leftarrow 1$
		\Else
			\State $edges[i] \leftarrow 0$
		\EndIf
	\EndFor
\EndFor
\Return $edges$
\end{algorithmic}
\end{algorithm}

@@

In this section we present a detailed description of Haralick's original work \cite{bb20239}.
The main idea of this algorithm is identical to that of the previous method: find zeros in 
the second derivative of the image. In this method, however, the input image is smoothly approximated through local bi-cubic
polynomial fitting. Then, when calculating the second derivative analytically, it is possible to find 
an equivalent expression to find the zeros of the second derivative of the polynomial as a function of 
its parameters.%\footnote{This implementation is slightly different from the traditional implementation 
%of the Haralick algorithm. We do not use the condition concerning the third derivative (to be negative), 
%in order to implement two edge detection methods using only the second derivative (along the 
%Marr-Hildreth algorithm).}

@@

Here, we present the interpolation method used in the original work of Haralick, although there are other options to do this \cite{getreuer}.

@@

\begin{equation*}
	\begin{array}{l}
		f_1 = f(x_1,y_1) = k_1 + k_2x_1 + k_3y_1 + k_4x_1^2 + k_5x_1y_1 + k_6y_1^2 + k_7x_1^3 + k_8x_1^2y_1 + k_9x_1y_1^2 + k_{10}y_1^3 \\
		f_2 = f(x_2,y_2) = k_1 + k_2x_2 + k_3y_2 + k_4x_2^2 + k_5x_2y_2 + k_6y_2^2 + k_7x_2^3 + k_8x_2^2y_2 + k_9x_2y_2^2 + k_{10}y_2^3 \\
		\vdots \\
		f_{25} = f(x_{25},y_{25}) = k_1 + k_2x_{25} + k_3y_{25} + k_4x_{25}^2 + k_5x_{25}y_{25} + k_6y_{25}^2 + k_7x_{25}^3 + k_8x_{25}^2y_{25} + k_9x_{25}y_{25}^2 + k_{10}y_{25}^3 . \\
	\end{array}
\end{equation*}
%
Using matrix notation, the system can be rewritten as
%
\begin{equation*}
	\begin{bmatrix} 
		f_1		\\ 
		f_2		\\ 
		\vdots	\\
		f_{25}
	\end{bmatrix} 
	= 
	\begin{bmatrix} 
		1 		& x_1 		& y_1 		& x_1^2 	& x_1y_1 		& y_1^2 	& \hdots 	& y_1^3 	\\
		1 		& x_2 		& y_2 		& x_2^2 	& x_2y_2 		& y_2^2 	& \hdots 	& y_2^3 	\\
		\vdots	& \vdots	& \vdots	& \vdots	& \vdots		& \vdots	& \ddots	& \vdots	\\
		1 		& x_{25}	& y_{25}	& x_{25}^2 	& x_{25}y_{25} 	& y_{25}^2 	& \hdots 	& y_{25}^3
	\end{bmatrix}
	\times
	\begin{bmatrix}
		k_1		\\
		k_2		\\
		\vdots	\\
		k_{10}
	\end{bmatrix}
	\Rightarrow \mathbf{f} = \mathbf{A}\mathbf{k} .\\
\end{equation*}
%
Then, the least squares problem can be solved by using the normal equations,
%
\begin{equation*}
	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k}.
\end{equation*}
%
Let us define a matrix $\mathbf{B} \in R^{10\times25}$ such that $\mathbf{B}=(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$, thus
%
$$ \mathbf{k} = \mathbf{B}\mathbf{f} $$
%
\begin{equation*}
	\begin{bmatrix}
		b_{1,1}		& b_{1,2}	& \hdots	& b_{1,25}	\\
		b_{2,1}		& b_{2,2}	& \hdots	& b_{2,25}	\\
		\vdots		& \vdots	& \ddots	& \vdots	\\
		b_{10,1}	& b_{10,2}	& \hdots	& b_{10,25}
	\end{bmatrix}
	\times
	\begin{bmatrix}
		f_1		\\
		f_2		\\
		\vdots	\\
		f_{25}
	\end{bmatrix}
	=
	\begin{bmatrix}
		k_1		\\
		k_2		\\
		\vdots	\\
		k_{10}
	\end{bmatrix} .\\
\end{equation*}

@@

The derivatives are obtained as follows.

@@

\begin{equation}
	f'''_{\theta}(\rho) = 6C_3 < 0 \ \ \Rightarrow \ \ C_3 < 0.
\end{equation}

By construction, $\theta$ is the direction of the gradient, which means that the edge will always be an ascending step. Thus by definition the function $f_{\theta}(\rho)$ is non-decreasing in a neighborhood of $\rho=0$, which means that $f'_{\theta}(0) \geq 0$. This in turn implies that $C_1 \geq 0$.

The second condition that the second derivative is equal to zero becomes

\begin{equation}\label{eq:haralick:cond_b_orig}
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0.	
\end{equation}

Ideally, if the edge point falls exactly on a grid location, the polynomial function 
should be zero at $\rho=0$. But due to the finite nature of the image grid, the edge could be located at non integer locations. Thus, the condition is relaxed asking that the polynomial becomes zero in a neighborhood of radius $\rho_0>0$ centered at $\rho=0$. 
With this relaxation, the condition \ref{eq:haralick:cond_b_orig} becomes

\begin{equation}
\exists\rho \in (-\rho_0,\rho_0) \:\:  \mathtt{such\:that}  \:\:  2C_2 + 6C_3\rho = 0 \\
\end{equation}

Thus, the solution is $\rho_*=-\frac{C_2}{3C_3} $ and imposing $|\rho_*|<\rho_0$ we obtain the second condition

\begin{equation}
\left| \frac{C_2}{3C_3} \right| < \rho_0,
\end{equation}

Finally,  the first condition is that the first derivative is non zero,

\begin{equation}\label{eq:cond:first}
f'_{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \neq 0 \\
\end{equation}

If we evaluate $f'_{\theta}$ on the solution $\rho_*$ using the above formula, we obtain

\begin{equation}\label{eq:cond:first:optim}
f'_{\theta}(\rho_*) = C_1 - \frac{C_2^2}{3C_3}  \\
\end{equation}

as $C_1>=0$ and $C_3<0$, $f'_{\theta}(\rho_*)$ will always be positive, thus the first condition always hold by construction and does not need to be verified.\\

@@

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\Require $im$ input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
\State Define the ten $5\times5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \Comment{See Table \ref{table:b_i}.}
\ForAll{pixel $i$ in image $im$}
	\State $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
	\For{$j=1$ \bf{to} $10$}
		\State $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
	\EndFor
	\State $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\State $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\If{{$|C_2/3C_3|\leq\rho_0$} \bf{and} {$C_3<0$}}
		\State $edges[i] \leftarrow 1$
	\Else
		\State $edges[i] \leftarrow 0$
	\EndIf
\EndFor
\Return $edges$
\end{algorithmic}
\end{algorithm}

@@

\begin{equation}
	LoG \stackrel{\triangle}{=}\nabla^2G(x,y)=\frac{\partial^2}{\partial^2 x}G(x,y) + \frac{\partial^2}{\partial^2 y}G(x,y)
\end{equation}

@@

As mentioned before, the size $n$ of the kernel and the standard deviation $\sigma$ of the exponential function 
are not independent. For the same $\sigma$ value , the function LoG has wider support than the
Gaussian function (i.e. LoG function has a slower decay), so a greater value of $n$ is needed to correcly sample the LoG kernel.\\

@@

Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm (including execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop. Note that the execution time of the algorithms Roberts, Prewitt and Sobel are the same, because they run simultaneously.}).

\begin{figure}[h!]
	\centering
	\subfigure[Roberts. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-a}\includegraphics[width=0.2\textwidth]{results/square127_roberts.png}}
	\quad
	\subfigure[Prewitt. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-b}\includegraphics[width=0.2\textwidth]{results/square127_prewitt.png}}
	\quad
	\subfigure[Sobel. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-c}\includegraphics[width=0.2\textwidth]{results/square127_sobel.png}}

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$. $\text{Exec-time}:0.03s$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$. $\text{Exec-time}:0.05s$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$. $\text{Exec-time}:0.04s$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image. The first three (\subref{fig:result1-a}, \subref{fig:result1-b} and \subref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\subref{fig:result1-d} and \subref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\subref{fig:result1-f}) corresponds to the Haralick algorithm. Note the rounded corners on the Marr-Hildreth results, due to Gaussian blur.}
	\label{fig:result1}
\end{figure}

In this simple case, the first derivative edge detection algorithms run better and faster. All algorithms show consistent results. Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm.

\begin{figure}[t!]
	\centering
	\includegraphics[width=0.48\textwidth]{results/molino_bw.png}
	\caption{Test image 2: Windmill (1000$\times$563 pixels).}
	\label{fig:original2}
\end{figure}

Now the performance of each algorithm is analyzed using a natural image (\textit{Windmill}) shown in Figure \ref{fig:original2}. 
The results of the first derivative edge detection algorithms are shown Figures \ref{fig:result3-b} to \ref{fig:result3-d}. Figures \ref{fig:result3-e} and \ref{fig:result3-f} show the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and Figure \ref{fig:result3-g} shows the results obtained using the Haralick algorithm. Table \ref{exectime2} summarizes the execution times for each of the algorithms.

\begin{table}[t!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.550 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $1.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $1.440 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.930 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including I/O) using the \emph{Windmill} image from Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
	\label{exectime2}
\end{table}

In this example the difference in performance between the first derivative and the second derivative algorithms is more meaningful.  For each image in Figure \ref{fig:result3}, we show an area of ​​interest, enlarging it to have a better view of the details. Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded). Neither the Haralick algorithm is able to detect it. However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel.

Another observation is that edges detected by Haralick's algorithm appear to be thicker than those detected with other ones. This may be due the regularity that Haralick assumes.

@@

Some observations: 
\begin{itemize}
	\item The first derivative algorithms, although they work properly and run fast, have some problems like discontinuity of the edges. They are also very sensitive to noise, which make them detect spurious edges (More sophisticated methods help to improve this, e.g. Canny edge detector \cite{Canny1986}). \\
	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results regarding edge details. This algorithm provides a first attempt to obtain regular edges, because of the filtering with a Gaussian kernel. This behavior can be seen in Lena's hair (Figure \ref{fig:example1}), or inside the oranges (example \ref{fig:example2}). \\
	\item As mentioned earlier, Haralick's algorithm is the first to assume greater regularity of the image in the neighborhood of a pixel (third order). This causes thicker edges, as shown in both examples. But, note that some edges are only detected with this algorithm, e.g. the lower edges of the oranges, which are in the shadow, in Figure \ref{fig:example2}. These edges are quite smooth, so that the Haralick model is well suited to them, and they do not represent an abrupt change in the intensity to be detected by other methods. \\
\end{itemize}

@@

The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since convolutions with larger kernels are performed. In real images, they have better behavior than first derivative algorithms.

>>>>

\newcommand{\demourl}{http://dev.ipol.im/~juanc/ipol_demo/sc_edges/}
\algrenewcommand\Return{\State \algorithmicreturn{} }%

@@

In the 70's, edge detection methods were implemented using small operators 
(such as Sobel masks), attempting to compute an approximation of the
first derivative of the image \cite{book:gonzalez:02:digital_image_processing}. Section \ref{sec:first} describes such algorithms 
and serves as an introduction to a more sophisticated 
analysis of the edge detection process.

% en este parrafo, puede ser cambiar trough por valley
In 1980 Marr and Hildreth \cite{segm:edge_region:marr:84:digital_step} argued that intensity changes are not independent 
of image scale, so edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or valley) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
Haralick's algorithm \cite{segm:edge_region:haralick:84:digital_step} is an alternative approach based on the second derivative 
which is also reviewed in Section \ref{sec:second}. This algorithm has the particularity of 
proposing a model to approximate the image locally around a point; then, using this model, 
an approximation to the second derivative of the image can be calculated analytically and 
finding edges is achieved by imposing a condition over the model parameters.

@@

@@

The edge detection methods discussed in the previous section are simply based on filtering the 
image with different masks, without taking into account the characteristics of the edges or 
the noise in the image.
The algorithms that will presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:84:digital_step} and Haralick \cite{segm:edge_region:haralick:84:digital_step}) 
are based on the second derivative of the image, and take steps to reduce noise before 
detecting edges in the image.

@@

\begin{algorithm}[t!]
\caption{Marr-Hildreth edge detection algorithm.}
\label{algo:mh}
\begin{algorithmic}[1]
\Require $im$ input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $t_{ZC}$.
\If{Gaussian kernel}
	\State $kernel \leftarrow$ generate\_gaussian\_kernel($n$,$\sigma$)
	\State $im_{smooth} \leftarrow$ convolution($im$,$kernel$)
	\State Define Laplacian operator $laplacian$ \Comment{See mask definition in Eq. \ref{eq:laplacian:mask}}
	\State $im_{lap} \leftarrow$ convolution($im_{smooth}$,$laplacian$)
\Else
\State $log\_kernel \leftarrow$ generate\_log\_kernel($n$,$\sigma$)
\State $im_{lap} \leftarrow$ convolution($im$,$log\_kernel$)
\EndIf
\State $max_L \leftarrow 0$
\ForAll{pixel $i$ in image $im_{lap}$}
	\If{$im_{lap}[i]>max_L$}
		\State $max_L \leftarrow im_{lap}[i]$
	\EndIf
\EndFor
\ForAll{pixel $i$ in image $im_{lap}$, except borders}
	\ForAll{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{lap}$}
		\If{($\mbox{sign}(im_{lap}[p_1])\neq\mbox{sign}(im_{lap}[p_2])$) \bf{and} ($|im_{lap}[p_1]-im_{lap}[p_2]|>th_{ZC}$)}
			\State $edges[i] \leftarrow 1$
		\Else
			\State $edges[i] \leftarrow 0$
		\EndIf
	\EndFor
\EndFor
\Return $edges$
\end{algorithmic}
\end{algorithm}

@@

In this section we present a detailed description of Haralick's original work \cite{segm:edge_region:haralick:84:digital_step}.
The main idea of this algorithm is identical to that of the previous method: find zeros in 
the second derivative of the image. In this method, however, the input image is smoothly approximated through local bi-cubic
polynomial fitting. Then, when calculating the second derivative analytically, it is possible to find 
an equivalent expression to find the zeros of the second derivative of the polynomial as a function of 
its parameters.%\footnote{This implementation is slightly different from the traditional implementation 
%of the Haralick algorithm. We do not use the condition concerning the third derivative (to be negative), 
%in order to implement two edge detection methods using only the second derivative (along the 
%Marr-Hildreth algorithm).}

@@

Here, we present the interpolation method used in the original work of Haralick, although there are other options to do this \cite{im_proc:getreuer:2011:linear_methods_for_image_interpolation}.

@@

\begin{equation*}
	\begin{array}{l}
		f_1 = f(x_1,y_1) = k_1 + k_2x_1 + k_3y_1 + k_4x_1^2 + k_5x_1y_1 + k_6y_1^2 + k_7x_1^3 + k_8x_1^2y_1 + k_9x_1y_1^2 + k_{10}y_1^3 \\
		f_2 = f(x_2,y_2) = k_1 + k_2x_2 + k_3y_2 + k_4x_2^2 + k_5x_2y_2 + k_6y_2^2 + k_7x_2^3 + k_8x_2^2y_2 + k_9x_2y_2^2 + k_{10}y_2^3 \\
		\vdots \\
		f_{25} = f(x_{25},y_{25}) = k_1 + k_2x_{25} + k_3y_{25} + k_4x_{25}^2 + k_5x_{25}y_{25} + k_6y_{25}^2 + k_7x_{25}^3 + k_8x_{25}^2y_{25} + k_9x_{25}y_{25}^2 + k_{10}y_{25}^3 . \\
	\end{array}
\end{equation*}
%
Using matrix notation, the system can be rewritten as
%
\begin{equation*}
	\begin{bmatrix} 
		f_1		\\ 
		f_2		\\ 
		\vdots	\\
		f_{25}
	\end{bmatrix} 
	= 
	\begin{bmatrix} 
		1 		& x_1 		& y_1 		& x_1^2 	& x_1y_1 		& y_1^2 	& \hdots 	& y_1^3 	\\
		1 		& x_2 		& y_2 		& x_2^2 	& x_2y_2 		& y_2^2 	& \hdots 	& y_2^3 	\\
		\vdots	& \vdots	& \vdots	& \vdots	& \vdots		& \vdots	& \ddots	& \vdots	\\
		1 		& x_{25}	& y_{25}	& x_{25}^2 	& x_{25}y_{25} 	& y_{25}^2 	& \hdots 	& y_{25}^3
	\end{bmatrix}
	\times
	\begin{bmatrix}
		k_1		\\
		k_2		\\
		\vdots	\\
		k_{10}
	\end{bmatrix}
	\Rightarrow \mathbf{f} = \mathbf{A}\mathbf{k} .\\
\end{equation*}
%
Then, the least squares problem can be solved by using the normal equations,
%
\begin{equation*}
	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k}.
\end{equation*}
%
Let us define a matrix $\mathbf{B} \in R^{10\times25}$ such that $\mathbf{B}=(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$, thus
%
$$ \mathbf{k} = \mathbf{B}\mathbf{f}. $$
%
\begin{equation*}
	\begin{bmatrix}
		b_{1,1}		& b_{1,2}	& \hdots	& b_{1,25}	\\
		b_{2,1}		& b_{2,2}	& \hdots	& b_{2,25}	\\
		\vdots		& \vdots	& \ddots	& \vdots	\\
		b_{10,1}	& b_{10,2}	& \hdots	& b_{10,25}
	\end{bmatrix}
	\times
	\begin{bmatrix}
		f_1		\\
		f_2		\\
		\vdots	\\
		f_{25}
	\end{bmatrix}
	=
	\begin{bmatrix}
		k_1		\\
		k_2		\\
		\vdots	\\
		k_{10}
	\end{bmatrix} .\\
\end{equation*}

@@

The derivatives are obtained as follows

@@

\begin{equation}
	f'''_{\theta}(\rho) = 6C_3 < 0 \ \ \Rightarrow \ \ C_3 < 0.
\end{equation}
%
The second condition that the second derivative is equal to zero becomes
%
\begin{equation}\label{eq:haralick:cond_b_orig}
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0.	
\end{equation}

Ideally, if the edge point falls exactly on a grid location, the polynomial function 
should be zero at $\rho=0$. But due to the finite nature of the image grid, the edge could be located at non integer locations. Thus, the condition is relaxed asking that the polynomial becomes zero in a neighborhood of radius $\rho_0>0$ centered at $\rho=0$. 
With this relaxation, the condition \ref{eq:haralick:cond_b_orig} becomes
%
\begin{equation}
\exists\rho \in (-\rho_0,\rho_0) \:\:  \mathtt{such\:that}  \:\:  2C_2 + 6C_3\rho = 0 .
\end{equation}
%
Thus, the solution is $\rho_*=-\frac{C_2}{3C_3} $ and imposing $|\rho_*|<\rho_0$ we obtain the second condition
%
\begin{equation}
\left| \frac{C_2}{3C_3} \right| < \rho_0.
\end{equation}
%
Finally,  the first condition is that the first derivative is non zero, 
%
\begin{equation}\label{eq:cond:first}
f'_{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \neq 0 \\
\end{equation}
%
If we evaluate $f'_{\theta}$ on the solution $\rho_*$ using the above formula, we obtain
%
\begin{equation}\label{eq:cond:first:optim}
f'_{\theta}(\rho_*) = C_1 - \frac{C_2^2}{3C_3}. 
\end{equation}
%
By construction, $\theta$ is the direction of the gradient, which means that the edge will always be an ascending step. Thus by definition the function $f_{\theta}(\rho)$ is non-decreasing in a neighborhood of $\rho=0$, which means that $f'_{\theta}(0) \geq 0$. This in turn implies that $C_1 \geq 0$.
As shown before, imposing the first condition implies that $C_3<0$, and we shown that $C_1>=0$, thus Equation \ref{eq:cond:first:optim} yields that $f'_{\theta}(\rho_*)$ will always be positive. This means that the first condition always hold by construction and does not need to be verified.\\

@@

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\Require $im$ input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
\State Define the ten $5\times5$ masks ($mask_1, \dots, mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \Comment{See Table \ref{table:b_i}.}
	\For{$j=1$ \bf{to} $10$} 
		\State $K_j \leftarrow$ convolution($neighbors$,$mask_j$) \Comment{$K_j$ is the convolution image of Eq. \ref{eq:convolution_b}.}
	\EndFor
\ForAll{pixel $i$ in image $im$}
	\For{$j=1$ \bf{to} $10$}
		\State $k_j \leftarrow$ $K_j(i)$ \Comment{$K_j(i)$ is the value of image $K_j$ at location $i$.}
	\EndFor
	\State $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\State $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\If{{$|C_2/3C_3|<\rho_0$} \bf{and} {$C_3<0$}}
		\State $edges[i] \leftarrow 1$
	\Else
		\State $edges[i] \leftarrow 0$
	\EndIf
\EndFor
\Return $edges$
\end{algorithmic}
\end{algorithm}

@@

\begin{equation}
	LoG \stackrel{\triangle}{=}\nabla^2G(x,y)=\frac{\partial^2}{\partial^2 x}G(x,y) + \frac{\partial^2}{\partial^2 y}G(x,y).
\end{equation}

@@

As mentioned before, the size $n$ of the kernel and the standard deviation $\sigma$ of the exponential function 
are not independent. For the same $\sigma$ value, the function LoG has wider support than the
Gaussian function (i.e. LoG function has a slower decay), so a greater value of $n$ is needed to correcly sample the LoG kernel.\\

@@

Figure \ref{fig:result1} shows the output of each algorithm including execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop. Note that the execution time of the algorithms Roberts, Prewitt and Sobel are the same, because they run simultaneously.}. In this simple case, the first derivative edge detection algorithms run better and faster. All algorithms show consistent results. Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm. 
%
\begin{figure}[h!]
	\centering
	\subfigure[Roberts. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-a}\includegraphics[width=0.2\textwidth]{results/square127_roberts.png}}
	\quad
	\subfigure[Prewitt. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-b}\includegraphics[width=0.2\textwidth]{results/square127_prewitt.png}}
	\quad
	\subfigure[Sobel. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-c}\includegraphics[width=0.2\textwidth]{results/square127_sobel.png}}

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$. $\text{Exec-time}:0.03s$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$. $\text{Exec-time}:0.05s$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$. $\text{Exec-time}:0.04s$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image. The first three (\subref{fig:result1-a}, \subref{fig:result1-b} and \subref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\subref{fig:result1-d} and \subref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\subref{fig:result1-f}) corresponds to the Haralick algorithm. Note the rounded corners on the Marr-Hildreth results, due to Gaussian blur.}
	\label{fig:result1}
\end{figure}
%
\begin{figure}[t!]
	\centering
	\includegraphics[width=0.48\textwidth]{results/molino_bw.png}
	\caption{Test image 2: Windmill (1000$\times$563 pixels).}
	\label{fig:original2}
\end{figure}

Now the performance of each algorithm is analyzed using a natural image (\textit{Windmill}) shown in Figure \ref{fig:original2}. 
The results of the first derivative edge detection algorithms are shown Figures \ref{fig:result3-b} to \ref{fig:result3-d}. Figures \ref{fig:result3-e} and \ref{fig:result3-f} show the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and Figure \ref{fig:result3-g} shows the results obtained using the Haralick algorithm. Table \ref{exectime2} summarizes the execution times for each of the algorithms. 
%
\begin{table}[t!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.550 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $1.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $1.440 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick											& $0.930 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including I/O) using the \emph{Windmill} image from Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
	\label{exectime2}
\end{table}
%
In this example the difference in performance between the first derivative and the second derivative algorithms is more meaningful.  For each image in Figure \ref{fig:result3}, we show an area of ​​interest, enlarging it to have a better view of the details. Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded). Neither the Haralick algorithm is able to detect it. However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel. Another observation is that edges detected by Haralick's algorithm appear to be thicker than those detected with other ones. This may be due the regularity that Haralick assumes.

@@

Some observations: 
\begin{itemize}
	\item The first derivative algorithms, although they work properly and run fast, have some problems like discontinuity of the edges. They are also very sensitive to noise, which make them detect spurious edges (More sophisticated methods help to improve this, e.g. Canny edge detector \cite{segm:edge_detection:canny:86:computational_edge}). \\
	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results regarding edge details. This algorithm provides a first attempt to obtain regular edges, because of the filtering with a Gaussian kernel. This behavior can be seen in Lena's hair (Figure \ref{fig:example1}), or inside the oranges (example \ref{fig:example2}). \\
	\item As mentioned earlier, Haralick's algorithm is the first to assume greater regularity of the image in the neighborhood of a pixel (third order). This causes thicker edges, as shown in both examples. But, note that some edges are only detected with this algorithm, e.g. the lower edges of the oranges, which are in the shadow, in Figure \ref{fig:example2}. These edges are quite smooth, so that the Haralick model is well suited to them, and they do not represent an abrupt change in the intensity to be detected by other methods. \\
\end{itemize}

@@

The second derivative algorithms (Marr-Hildreth \& Haralick) involve several more operations, since convolutions with larger kernels are performed. In real images, they have better behavior than first derivative algorithms.
